---
title: "The Effect on Vitamin C on Guinea Pigs' Tooth Growth"
author: "Giulio Mario Martena"
date: "6/4/2020"
output: pdf_document
---
\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.path='Figs/', warning=FALSE, message=FALSE)
```

## Overview
The general goal of this analysis is to analyze the `ToothGrowth` dataset, part of the `datasets` library preincluded in R. From the documentation of the dataset, we learn that among 60 guinea pigs, *each animal received one of three dose levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods, **orange juice** or **ascorbic acid** (a form of vitamin C and coded as VC)*.

```{r, echo=FALSE}
require(datasets); require(ggplot2); require(dplyr)
```
```{r}
data("ToothGrowth")
ToothGrowth$dose <- factor(ToothGrowth$dose)
ToothGrowth<-as_tibble(ToothGrowth);head(ToothGrowth)
```

## Exploratory Analysis

From *figure 1* (cfr. Appendix session below) it appears that a clear separation between the two treatments is only visible with a dose of 1 mg/day; for 2 mg/day the separation is not clear, even though there it appears that ascorbic acid has a bigger variance.

## Hypothesis Testing  

First of all, I wonder if mean from the tooth growth is the same **globally** and **divided by dosage**.

* *I will assume that the variance is the same and that they used 6 different sets of guinea pigs in order not to have results affected by previous treatments.*

### Hypothesis 1: no dosage separation   
$$H_0: \mu_{OJ} = \mu_{VC}$$ $$H_1: \mu_{OJ} \neq \mu_{VC}$$

```{r}
t.test(len ~ supp, ToothGrowth, paired = FALSE, var.equal = TRUE)
```
Since $p > 0.05$ and the confidence interval includes 0, so even though $\mu_{OJ}>\mu_{VC}$, $H_0$ *cannot* be rejected; hence, globally, we can assume that the means are in fact equal.

### Hypothesis 2: 0.5mg/day dosage

$$H_0: \mu_{OJ,.5} = \mu_{VC,.5}$$ $$H_1: \mu_{OJ,.5} \neq \mu_{VC,.5}$$

```{r}
t.test(len ~ supp, data = filter(ToothGrowth, dose == 0.5), paired = FALSE, var.equal = TRUE)
```
The p-value is smaller than .05, and the confidence interval does not include 0. In this case we can reject $H_0$ and accept that $\mu_{OJ,.5}>\mu_{VC,.5}$, which means that for a dosage of 0.5mg/day the orange juice causes more teeth growth than ascorbic acid.

### Hypothesis 3: 1mg/day dosage

$$H_0: \mu_{OJ,1} = \mu_{VC,1}$$ $$H_1: \mu_{OJ,1} \neq \mu_{VC,1}$$

```{r}
t.test(len ~ supp, data = filter(ToothGrowth, dose == 1), paired = FALSE, var.equal = TRUE)
```
Just as in hypothesis 2, T-Test suggests that $p<0.05$ and the confidence interval does not include 0. $H_0$ is rejected, hence for a dosage of 1mg/day it appears that orange juice grows teeth more than its rival.

### Hypothesis 4: 2mg/day dosage

$$H_0: \mu_{OJ,2} = \mu_{VC,2}$$ $$H_1: \mu_{OJ,2} \neq \mu_{VC,2}$$

```{r}
t.test(len ~ supp, data = filter(ToothGrowth, dose == 2), paired = FALSE, var.equal = TRUE)
```
Since $p > 0.05$ and the confidence interval includes 0, so even though $\mu_{OJ,2}>\mu_{VC,2}$, $H_0$ *cannot* be rejected; hence, globally, we can assume that the both supplements bring the same growth for a 2mg/day dosage.

From the boxplot it would also appear a similar growth for 1mg/day dosage of orange juice and 2mg/day dosage of ascorbic acid. Is it true?

### Hypothesis 5: 1mg/day of orange juice vs 2mg/day of ascorbic acid

$$H_0: \mu_{OJ,1} = \mu_{VC,2}$$ $$H_1: \mu_{OJ,1} \neq \mu_{VC,2}$$

```{r}
h <- bind_rows(ToothGrowth %>%
                  filter((dose==1 & supp=='OJ') | (dose==2 & supp=='VC'))
                  )
t.test(len ~ supp, h, paired = FALSE, var.equal = TRUE)
```
The $p>0.05$ and the C.I. includes 0. The $H_0$ cannot be rejected, which means that 2 mg/day of ascorbic acid seems to affect the teeth as much 1 mg/day of orange juice.

A similar pattern also appears in 1mg/day of orange juice and 2mg/day of the same supplement.

### Hypothesis 6: 1mg/day vs 2mg/day of orange juice

$$H_0: \mu_{OJ,2} = \mu_{OJ,1}$$ $$H_1: \mu_{OJ,2} \neq \mu_{OJ,1}$$

```{r}
h <- bind_rows(ToothGrowth %>%
                  filter((dose== c(1,2) & supp=='OJ'))
                  )
t.test(len ~ dose, h, paired = FALSE, var.equal = TRUE)
```
$p < 0.05$ and the confidence interval does not include 0 (which would be the difference in means if they were equal), hence $H_0$ is rejected. 2mg/day of orange juice appear to bring more tooth growth to the table than half that dosage.

## Conclusion
The teeth length seems to **grow with higher dosage of vitamin C**. Furthermore, orange juice appears to influence it more significantly on lower dosages, whereas on 2mg/day dosage the effect is pretty much similar (cfr. hypothesis 4).

## Appendix


```{r, fig.cap="Tooth Length vs Supplement and Dose Levels",echo=FALSE}
ggplot(ToothGrowth, aes(x=dose, y=len)) + geom_boxplot(aes(fill = supp)) + labs(x="Dose[mg/day]",y="Tooth Length")
```
