---
title: "My Fancy Presentation"
author: "Giulio Martena"
date: "7/26/2020"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## My new slide
Some content

Plain text

*italicized*  
**bold**  
`some code`  

## My new slide

```{r,comment='', echo=TRUE, eval=FALSE}
head(mtcars)
```

## Another slide
```{r, fig.align='center', fig.cap='My awesome figure'}
plot(mtcars$wt, mtcars$mpg, xlab='Weight',ylab='MPG')
```

***
A slide without a title

## My title

### A top Level subheading

#### A top second

- Bullet 1
- Bullet 2

1. Item a
2. Item b

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

## My first map

```{r}
library(leaflet)
leaflet() %>% addTiles()
```

## Duomo di Lecce

```{r}
my_map <- leaflet() %>%
    addTiles() %>%
    addMarkers(lat = 40.35, lng = 18.167, popup = 'Duomo di Lecce')
my_map
```


## Random markers

```{r}
df <- data.frame(lat=runif(20, min = 39.2, max=39.3),
                 lng=runif(20, min = -76.6,max = -76.5))

df %>%
    leaflet() %>%
    addTiles() %>%
    addMarkers()
```